Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-05-05 paper.qmd
Updated the projection function

```{r}
#| label: langevitour-best-fit-two-curvy-proj-tsne

@@

#nldr_two_curvy2 + 
  trimesh_two_curvy_removed1_tsne_with_data +
  #two_curvy_proj_tsne_first_model1 +
  #two_curvy_proj_tsne_all_model1 +
  plot_layout(ncol = 3)
```

@@

names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")

model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")

@@

## Compute 2D distances
distance_two_curvy4 <- cal_2d_dist(
  tr_coord_df = tr_from_to_df_two_curvy4, 
  start_x = "x_from", 
  start_y = "y_from", 
  end_x = "x_to", 
  end_y = "y_to", 
  select_vars = c("from", "to", "distance"))

## To find the benchmark value
benchmark_two_curvy4 <- find_lg_benchmark(
  distance_edges = distance_two_curvy4, 
  distance_col = "distance")

benchmark_two_curvy4 <- 0.2

## Benchmark 1

distance_df_small_edges_two_curvy4 <- distance_two_curvy4 |>
  filter(distance < benchmark_two_curvy4) |> ## 0.221
  filter(!(row_number() %in% c(35, 89, 83, 25, 80, 56, 13, 59, 16)))

tr_from_to_df_two_curvy4 <- inner_join(
  tr_from_to_df_two_curvy4, distance_df_small_edges_two_curvy4,
  by = c("from", "to"))

@@

model_df <- dplyr::left_join(
  distance_df_small_edges_two_curvy4 |> select(-distance), 
  projected_model_df, 
  by = c("from" = "ID"))

names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")

model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")

@@

tsne_pbmc_scaled_best <- algo_obj_pbmc$nldr_scaled
distance_pbmc <- algo_obj_pbmc$distance_df
tr_from_to_df_pbmc <- algo_obj_pbmc$tr_from_to_df
benchmark_pbmc <- algo_obj_pbmc$benchmark
df_bin_centroids_pbmc <- algo_obj_pbmc$df_bin_centroids
df_bin_pbmc <- algo_obj_pbmc$df_bin

distance_df_small_edges_pbmc <- distance_pbmc |>
  filter(distance < benchmark_pbmc) #benchmark_pbmc

# tr_from_to_df_pbmc <- tr_from_to_df_pbmc |>
#   mutate(ID = row_number()) |>
#   filter(!(ID %in% c(19, 20, 23, 24, 29, 33, 48,
#                      183, 181, 178, 176))) #19

tr_from_to_df_pbmc <- right_join(
  tr_from_to_df_pbmc, distance_df_small_edges_pbmc,
  by = c("from", "to")) #|>
  #select(-ID)

@@

langevitour::langevitour(df_model_data_pbmc_n[1:(length(df_model_data_pbmc_n)-1)],
                         lineFrom = distance_df_small_edges_pbmc$from,
                         lineTo = distance_df_small_edges_pbmc$to,
                         group = factor(df_model_data_pbmc_n$type,
                                        c("cluster1", "cluster2", "cluster3", "model")),
                         levelColors = c("#8dd3c7", "#fdb462", "#bebada", "#000000"))
```

@@

proj_obj1 <- get_projection(projection = projection, 
                            proj_scale = 1.23, 
                            scaled_data = scaled_pbmc_data, 
                            scaled_data_model = scaled_pbmc_data_model, 
                            distance_df_small_edges = distance_df_small_edges_pbmc, 
                            axis_param = list(limits = 1, 
                                              axis_scaled = 1,
                                              axis_pos_x = -0.48, 
                                              axis_pos_y = -0.48, 
                                              threshold = 0.082))

@@

proj_obj2 <- get_projection(projection = projection, 
                            proj_scale = 1.25, 
                            scaled_data = scaled_pbmc_data, 
                            scaled_data_model = scaled_pbmc_data_model, 
                            distance_df_small_edges = distance_df_small_edges_pbmc, 
                            axis_param = list(limits = 1.35, 
                                              axis_scaled = 1,
                                              axis_pos_x = -0.75, 
                                              axis_pos_y = -0.75, 
                                              threshold = 0.12))

>>>>

```{r}
#| label: langevitour-best-fit-two-curvy-proj-tsne
#| eval: false

@@

#nldr_two_curvy2 + 
  trimesh_two_curvy_removed1_tsne_with_data +
  two_curvy_proj_tsne_first_model1 +
  two_curvy_proj_tsne_all_model1 +
  plot_layout(ncol = 3)
```

@@

names(model_df)[7:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")

model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
names(model_df)[(6 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")

@@

@@

names(model_df)[7:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")

model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
names(model_df)[(6 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")

@@

tsne_pbmc_scaled_best <- algo_obj_pbmc$nldr_scaled
distance_pbmc <- algo_obj_pbmc$distance_df
tr_from_to_df_pbmc <- algo_obj_pbmc$tr_from_to_df
df_bin_centroids_pbmc <- algo_obj_pbmc$df_bin_centroids
df_bin_pbmc <- algo_obj_pbmc$df_bin

@@

@@

@@
